---
title: "Mini Project 2"
author: "Entong Li, Zijia Cao, Jingyu Xie, Xindi Lu"
date: "10/31/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
load("CCES16_Common_OUTPUT_Feb2018_VV.RData")
library(gapminder)
library(dplyr)
library(tidyverse)
library(GGally)
library(gridExtra)
cb_palette = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
cbp <- function(x) scale_color_manual(values = cb_palette)
variables <- c("CC16_326", "CC16_410a", "commonweight_vv_post", "tookpost","gender","educ","race","pid7", "CC16_331_1","CC16_331_2","CC16_331_3","CC16_331_7")
```

# 1. Creating the data frame used in exploration:

```{r}
obama <- x[variables] %>%
  filter(CC16_326 == "Barack Obama", tookpost == "Yes")
obama$vote.T <- ifelse(obama$CC16_410a == "Donald Trump (Republican)", 1,0)

#For CC16_331_1 and CC16_331_3, "Yes" is a pro-immigration response.
for (i in c(9,11)){
  obama[i] <- ifelse(obama[i] == "Yes", 1 , 0)
}

#For CC16_331_2 and CC16_331_7, "No" is a pro-immigration response.
for (i in c(10,12)){
  obama[i] <- ifelse(obama[i] == "No", 1 , 0)
}

#Level of pro-immigration: 4 is the most, 0 is the lease.
obama$immig.at <- as.numeric(obama$CC16_331_1 + obama$CC16_331_2 + obama$CC16_331_3 + obama$CC16_331_7)

#Recode the race to four levels.
obama$race <- recode_factor(obama$race,"White" = "White", "Black" = "Black", "Hispanic" = "Hispanic", .default = "Other")

#Recode the party identification to numeric variables:
#Strong to Lean Democrat: 3 to 1;
#Strong to Lean Republican: -3 to -1;
#Independent and others: 0
obama$pid7 <- recode_factor(obama$pid7, "Strong Democrat" = "Strong Democrat", "Not very strong Democrat" = "Not very strong Democrat", "Lean Democrat" = "Lean Democrat", "Strong Republican" = "Strong Republican", "Not very strong Republican" = "Not very strong Republican", "Lean Republican" = "Lean Republican", .default = "Other")

#Recode the education level to numeric variables:
#The higer value means higher education level, from 1-6.
#Skipped and Not Asked will be 0.
obama$educ <- recode(obama$educ, "No HS" = 1, "High school graduate" = 2, "Some college" = 3, "2-year" = 4, "4-year" = 5, "Post-grad" = 6, .default = 0)


#Recode the gender into three groups: Male, Female, Unknown(Skipped and Not asked)
obama$gender <- recode_factor(obama$gender, "Male" = "Male", "Female" = "Female", .default = "Unknown")
```


# 2.

### Immigration attitude with gender:

```{r}

our.logit = function(x){
  coe = coef(obama.weighted.logit)
  y = coe[1] + coe[2] * x
  return(exp(y) / (1 + exp(y)))
}

#Immigration attitude with gender:
immig.gender.1 = glm(vote.T ~ immig.at + gender, family = quasibinomial, weights = commonweight_vv_post, data = obama)
immig.gender.2 = glm(vote.T ~ immig.at * gender, family = quasibinomial, weights = commonweight_vv_post, data = obama)
gender.df.1 = expand.grid(immig.at = 0:4, gender = factor(c("Male","Female")))
gender.pred.1 = predict(immig.gender.1, type = "response", newdata =gender.df.1 )
gender.pred.df.1 = data.frame(gender.df.1, vote.T.prob = as.vector(gender.pred.1))
gender.1<- ggplot(gender.pred.df.1, aes(x = immig.at, y = vote.T.prob, group = gender, color = gender)) + geom_line() + xlab("Immigration attitude") + ylab("Probability of voting Donald Trump") + ggtitle("Without interaction with gender")
gender.df.2 = expand.grid(immig.at = 0:4, gender = factor(c("Male","Female")))
gender.pred.2 = predict(immig.gender.2, type = "response", newdata =gender.df.2 )
gender.pred.df.2 = data.frame(gender.df.2, vote.T.prob = as.vector(gender.pred.2))
gender.2 <- ggplot(gender.pred.df.2, aes(x = immig.at, y = vote.T.prob, group = gender, color = gender)) + geom_line() + xlab("Immigration attitude") + ylab("Probability of voting Donald Trump")+ggtitle("With interaction with gender")
grid.arrange(gender.1,gender.2, nrow = 1)
```

The immigration attitude doesn't interact with gender. Because the slopes for male and female from the model without interaction with gender do not have difference with the slopes of male and female from the model with interaction with gender.  \n


### Immigration attitude with education level:

```{r}
#Immigration attitude with education level:
immig.educ.1 = glm(vote.T ~ immig.at + educ, family = quasibinomial, weights = commonweight_vv_post, data = obama)
immig.educ.2 = glm(vote.T ~ immig.at * educ, family = quasibinomial, weights = commonweight_vv_post, data = obama)
educ.df.2 = expand.grid(immig.at = 0:4, educ = 1:6)
educ.pred.2 = predict(immig.educ.2, type = "response", newdata =educ.df.2 )
educ.pred.df.2 = data.frame(educ.df.2, vote.T.prob = as.vector(educ.pred.2))
educ.2 <- ggplot(educ.pred.df.2, aes(x = immig.at, y = vote.T.prob, group = educ, color = educ)) + geom_line() + xlab("Immigration attitude") + ylab("Probability of voting Donald Trump") + labs(color = "education level") + ggtitle("With interaction with\neducation level")
educ.df.1 = expand.grid(immig.at = 0:4, educ = 1:6)
educ.pred.1 = predict(immig.educ.1, type = "response", newdata =educ.df.1 )
educ.pred.df.1 = data.frame(educ.df.1, vote.T.prob = as.vector(educ.pred.1))
educ.1<- ggplot(educ.pred.df.1, aes(x = immig.at, y = vote.T.prob, group = educ, color = educ)) + geom_line() + xlab("Immigration attitude") + ylab("Probability of voting Donald Trump") + labs(color = "education level")+ ggtitle("Without interaction with\neducation level")
grid.arrange(educ.1,educ.2, nrow = 1)
```

The immigration attitudes interact with educational level. Because for the model without interaction with education level, each level of education level seems to have similar, almost the same, slope, and when the immigration attitude is the same, higher level of education will have higher probability of voting Donald Trump. In contrast, for the model with interaction with education level, different education level will have different slope, and all of them will intersect when immigration attitude is around 0.7. As the education level will affect the probability of voting Donald Trump differently when there is an interaction between education level and immigration attitude, the interaction with education level should be included in model. \n


### Immigration attitude with race:

```{r}
#Immigration attitude with race:
immig.race.1 = glm(vote.T ~ immig.at + race, family = quasibinomial, weights = commonweight_vv_post, data = obama)
immig.race.2 = glm(vote.T ~ immig.at * race , family = quasibinomial, weights = commonweight_vv_post, data = obama)
race.df.2 = expand.grid(immig.at = 0:4, race = factor(c("White","Black","Hispanic","Other")))
race.pred.2 = predict(immig.race.2, type = "response", newdata =race.df.2 )
race.pred.df.2 = data.frame(race.df.2, vote.T.prob = as.vector(race.pred.2))
race.2<- ggplot(race.pred.df.2, aes(x = immig.at, y = vote.T.prob, group = race, color = race)) + geom_line() + xlab("Immigration attitude") + ylab("Probability of voting Donald Trump") + ggtitle("With interaction with race")
race.df.1 = expand.grid(immig.at = 0:4, race = factor(c("White","Black","Hispanic","Other")))
race.pred.1 = predict(immig.race.1, type = "response", newdata =race.df.1 )
race.pred.df.1 = data.frame(race.df.1, vote.T.prob = as.vector(race.pred.1))
race.1 <- ggplot(race.pred.df.1, aes(x = immig.at, y = vote.T.prob, group = race, color = race)) + geom_line() + xlab("Immigration attitude") + ylab("Probability of voting Donald Trump") + ggtitle("Without interaction with race")
grid.arrange(race.1,race.2, nrow = 1)
```

The immigration attitude interacts with the race. Because in model includes the interaction with race, Hispanic, Other, and White seem to have different slope from  in the model does not include the interaction with race. Thus the interaction between immigration attitude and race should be included. \n


### Immigration attitude with party:

```{r}
#Immigration attitude with party:
immig.party.1 = glm(vote.T ~ immig.at + pid7, family = quasibinomial, weights = commonweight_vv_post, data = obama)
immig.party.2 = glm(vote.T ~ immig.at * pid7 , family = quasibinomial, weights = commonweight_vv_post, data = obama)
# summary(immig.party.1)
# summary(immig.party.2)
party.df.2 = expand.grid(immig.at = 0:4, pid7 = factor(c("Strong Democrat", "Not very strong Democrat", "Lean Democrat", "Strong Republican",  "Not very strong Republican", "Lean Republican",  "Other")))
party.pred.2 = predict(immig.party.2, type = "response", newdata =party.df.2 )
party.pred.df.2 = data.frame(party.df.2, vote.T.prob = as.vector(party.pred.2))
party.2 <- ggplot(party.pred.df.2, aes(x = immig.at, y = vote.T.prob, group = pid7, color = pid7)) + geom_line() + xlab("Immigration attitude") + ylab("Probability of voting Donald Trump") + labs(color = "Party")
party.df.1 = expand.grid(immig.at = 0:4, pid7 = factor(c("Strong Democrat", "Not very strong Democrat", "Lean Democrat", "Strong Republican",  "Not very strong Republican", "Lean Republican",  "Other")))
party.pred.1 = predict(immig.party.1, type = "response", newdata =party.df.1 )
party.pred.df.1 = data.frame(party.df.1, vote.T.prob = as.vector(party.pred.1))
party.1<- ggplot(party.pred.df.1, aes(x = immig.at, y = vote.T.prob, group = pid7, color = pid7)) + geom_line() + xlab("Immigration attitude") + ylab("Probability of voting Donald Trump") + labs(color = "Party") 
party.1 + ggtitle("Without interaction with party")
party.2 + ggtitle("With interaction with party")
```


The immigration attitude interacts with party.  In the model with interaction with party, Lean Democrat, Lean Republican, Not very strong Republican, and Strong Democrat have different slope from in the model without interaction with party. Thus the interaction between immigration attitude and party should be included.  \n


#3


```{r}
#creating a test dataset

# test<-obama
# test$race2 <- recode(obama$race, White = "White", Black = "Black", Hispanic = "Hispanic", .default = "Other") 
# test$educ2 <- as.numeric(test$educ)
# test$pid72 <- as.numeric(test$pid7)
# test$gender2 <- as.numeric(test$gender)
# test$race2 <- as.numeric(test$race2)
# test$switching <- ifelse(obama$CC16_326 == 'Barack Obama' & obama$CC16_410a == 'Donald Trump (Republican)', 1, 0)
# test <- test[!is.na(test$switching), ]

```


```{r}
library(arm)
model1 <- glm(vote.T ~ race + pid7 + educ , family = 'binomial', data = obama, weights = commonweight_vv_post)
display(model1)

model2 <- glm(vote.T ~ immig.at + race + pid7 + educ + gender + immig.at:race + immig.at:educ + immig.at:pid7, family = 'binomial', data = obama, weights = commonweight_vv_post)
display(model2)

# include interactions as necessary
# model2.1 <- glm(switching ~ immig.at + race2 + pid72 + educ2 + gender2 + immig.at * race2 + immig.at * educ2, family = binomial(link='logit'), data = test, weights = commonweight_vv_post)
# summary(model2.1)

```


```{r}
#1. With immigration attidue:

with.img = expand.grid(immig.at = 0:4, pid7 = factor(c("Strong Democrat", "Not very strong Democrat", "Lean Democrat", "Strong Republican",  "Not very strong Republican", "Lean Republican",  "Other")), race = factor(c("White","Black","Hispanic","Other")), educ = 1:6, gender = factor(c("Male","Female")))
with.img.pred = predict(model2, type = "response", newdata = with.img )
with.img.pred.df = data.frame(with.img, vote.T.prob = as.vector(with.img.pred))

#2. Without immigration attitude:

wout.img = expand.grid(pid7 = factor(c("Strong Democrat", "Not very strong Democrat", "Lean Democrat", "Strong Republican",  "Not very strong Republican", "Lean Republican",  "Other")), race = factor(c("White","Black","Hispanic","Other")), educ = 1:6, gender = factor(c("Male","Female")))
wout.img.pred = predict(model1, type = "response", newdata = wout.img)
wout.img.pred.df = data.frame(wout.img, vote.T.prob = as.vector(wout.img.pred))
```


```{r}
gender.with <- ggplot(data = with.img.pred.df, aes(y = vote.T.prob, x = gender)) + geom_boxplot() + ggtitle("With immigration attitude") + ylab("Probability of voting to Donald Trump")
gender.wout <- ggplot(data = wout.img.pred.df, aes(y = vote.T.prob, x = gender)) + geom_boxplot() + ggtitle("Without immigration attitude") + ylab("Probability of voting to Donald Trump")
grid.arrange(gender.with, gender.wout, nrow = 1)

race.with <- ggplot(data = with.img.pred.df, aes(y = vote.T.prob, x = race)) + geom_boxplot() + ggtitle("With immigration attitude") + ylab("Probability of voting to Donald Trump")
race.wout <- ggplot(data = wout.img.pred.df, aes(y = vote.T.prob, x = race)) + geom_boxplot() + ggtitle("Without immigration attitude") + ylab("Probability of voting to Donald Trump")
grid.arrange(race.with, race.wout, nrow = 1)

educ.with <- ggplot(data = with.img.pred.df, aes(y = vote.T.prob, x = educ, group = educ, color = educ)) + geom_boxplot() + ggtitle("With immigration attitude") + ylab("Probability of voting to Donald Trump")
educ.wout <- ggplot(data = wout.img.pred.df, aes(y = vote.T.prob, x = educ, group = educ, color = educ)) + geom_boxplot() + ggtitle("Without immigration attitude") + ylab("Probability of voting to Donald Trump")
grid.arrange(educ.with, educ.wout, nrow = 1)

party.with <- ggplot(data = with.img.pred.df, aes(y = vote.T.prob, x = pid7)) + geom_boxplot() + ggtitle("With immigration attitude") + ylab("Probability of voting to Donald Trump")
party.wout <- ggplot(data = wout.img.pred.df, aes(y = vote.T.prob, x = pid7)) + geom_boxplot() + ggtitle("Without immigration attitude") + ylab("Probability of voting to Donald Trump")
grid.arrange(party.with, party.wout, nrow = 1)
```



```{r}
# group1 <- data.frame(race2 = 1, pid72 = 7, educ2 = 1, gender2 = 1, immig.at = 0) # white strong republican no high school male 
# predict(model1, newdata = group1, type = "response")
# predict(model2, newdata = group1, type = "response")
# predict(model2.1, newdata = group1, type = "response")
# 
# group2 <- data.frame(race2 = 3, pid72 = 1, educ2 = 6, gender2 = 2, immig.at = 4) 
# hispanic， strong democrat， Post-grad， female 
# predict(model1, newdata = group2, type = "response")
# predict(model2, newdata = group2, type = "response")
# predict(model2.1, newdata = group2, type = "response")
```

